\bf What is the Best Way to Display Two Time Series to Read Association?

### 2016-02-09 paper/paper.Rnw
editing section 4

\begin{figure}[htbp]
 \centering
<<results, dependson='setup', fig.width=6, fig.height=8, out.width='0.7\\textwidth', echo=FALSE, warning=FALSE, error=FALSE, message=FALSE>>=
ggplot(data=bylineup, aes(x=r, y=percent, colour=design)) + geom_smooth(se=F) + geom_point() + facet_grid(n~smoothed) + 
  scale_colour_discrete(name = "Plot Design") + labs(y="Detection rate", x="Correlation") +
  theme(legend.position="bottom")
@
  \caption{Response accuracy for unsmoothed data, broken by unsmoothed (left), smoothed (right) and sample size. The dots are data points, the lines represent a naive model average fit. Both plot designs have similar performance for positively correlated data, however the overlaid line graphs are much weaker at revealing negative correlation.}
    \label{results}
\end{figure}

@@

\begin{figure}[htbp]
  \centering
<<powerdiff, dependson='setup', fig.width=6, fig.height=8, out.width='0.7\\textwidth', echo=FALSE, error=FALSE, message=FALSE, warning=FALSE>>=
prop <- select(bylineup, smoothed, n, evals, total, r, design)
prop <- summarise(group_by(prop, smoothed, n, r), l.attempt=sum(ifelse(design=="Overlaid Lines", evals, 0)),
                                                  l.detect=sum(ifelse(design=="Overlaid Lines", total, 0)),
                                                  s.attempt=sum(ifelse(design=="Scatterplot", evals, 0)),
                                                  s.detect=sum(ifelse(design=="Scatterplot", total, 0)))
alpha <- 1-(1-0.95^(1/48))/2 
##8 correlations * 6 design = 48 boxplots, so this the students t quantile for the joint 95% CI
prop <- mutate(prop, l.power = l.detect/l.attempt, 
               s.power = s.detect/s.attempt,
               df = s.attempt + l.attempt - 1, 
               difference=s.power-l.power, 
               t = qt(alpha, df),
               se = sqrt(s.power*(1-s.power)/s.attempt+l.power*(1-l.power)/l.attempt),
               upper = difference + t*se,
               lower = difference - t*se)
ggplot(data=prop) + geom_errorbar(aes(x=r, ymax=upper, ymin=lower)) + geom_point(aes(x=r, y=difference)) +
        geom_hline(aes(yintercept=0, colour="red")) + labs(x="Correlation", y="Power difference") + facet_grid(n~smoothed)
@
 \caption{The difference in estimated power (scatter plot power minus overlaid line power) to reveal the true plot by smoothness and sample size with a joint 95\% confidence interval. The red line indicates the point where both plot designs are equally powerful The largest differences appear in negative correlation, where the overlaid lines power drops substantially.}
 \label{powerdiff}
\end{figure}

\begin{figure}[htbp]
  \centering
<<samplesize, dependson='setup', fig.width=6, fig.height=8, out.width='0.7\\textwidth', echo=FALSE, error=FALSE, message=FALSE, warning=FALSE>>=
ggplot(data=subset(bylineup, smoothed==0), aes(x=r, y=percent, colour=factor(n))) + geom_smooth(se=F) + geom_point() + facet_grid(design~.) + scale_colour_discrete(name = "Sample size") +  labs(y="Detection rate", x="Correlation") + theme(legend.position="bottom")
@
 \caption{Replotting figure~\ref{results} to highlight the effects of changes in sample size of unsmoothed data. The scatter plot appears to universally benefit from increased sample sizes but the overlaid lines only see a benefit in positive correlation.}
 \label{samplesize}
\end{figure}

@@

%The correlation:sign interaction shows that the scatter plot has close to symmetric performance for positive (correlation:sign + correlation) and negative (-correlation) correlation while the overlaid line graph is substantially weaker for negative values of correlation.
Smoothing out the excess 'noise' from a time series had a minimal effect on scatter plots, demonstrated by $\beta_{4} + \beta_{10}$, but it did have a significant positive impact on the power of overlaid line graphs as smoothing induces a large visual difference (See Figure~\ref{smoothing}. Increasing the sample size, $t$, improved the power of the scatter plot but it was not statistically significant for the overlaid line graph.

>>>>

\begin{figure}[htbp]
 \centering
<<results, dependson='setup', fig.width=6, fig.height=8, out.width='0.7\\textwidth', echo=FALSE, warning=FALSE, error=FALSE, message=FALSE>>=
ggplot(data=bylineup, aes(x=r, y=percent, colour=design)) + geom_smooth(se=F) + geom_point() + facet_grid(n~smoothed) + 
  scale_colour_discrete(name = "Plot Design") + labs(y="Detection rate", x="Correlation") +
  theme(legend.position="bottom")
@
  \caption{Response accuracy for unsmoothed data, broken by unsmoothed (left), smoothed (right) and sample size. The dots are data points, the lines represent a naive model average fit. Both plot designs have similar performance for positively correlated data, however the overlaid line graphs are much weaker at displaying negative correlation.}
    \label{results}
\end{figure}

@@

\begin{figure}[htbp]
  \centering
<<powerdiff, dependson='setup', fig.width=6, fig.height=8, out.width='0.7\\textwidth', echo=FALSE, error=FALSE, message=FALSE, warning=FALSE>>=
prop <- select(bylineup, smoothed, n, evals, total, r, design)
prop <- summarise(group_by(prop, smoothed, n, r), l.attempt=sum(ifelse(design=="Overlaid Lines", evals, 0)),
                                                  l.detect=sum(ifelse(design=="Overlaid Lines", total, 0)),
                                                  s.attempt=sum(ifelse(design=="Scatterplot", evals, 0)),
                                                  s.detect=sum(ifelse(design=="Scatterplot", total, 0)))
alpha <- 1-(1-0.95^(1/48))/2 
##8 correlations * 6 design = 48 boxplots, so this the students t quantile for the joint 95% CI
prop <- mutate(prop, l.power = l.detect/l.attempt, 
               s.power = s.detect/s.attempt,
               df = s.attempt + l.attempt - 1, 
               difference=s.power-l.power, 
               t = qt(alpha, df),
               se = sqrt(s.power*(1-s.power)/s.attempt+l.power*(1-l.power)/l.attempt),
               upper = difference + t*se,
               lower = difference - t*se)
ggplot(data=prop) + geom_errorbar(aes(x=r, ymax=upper, ymin=lower)) + geom_point(aes(x=r, y=difference)) +
        geom_hline(aes(yintercept=0, colour="red")) + labs(x="Correlation", y="Power difference") + facet_grid(n~smoothed)
@
 \caption{The difference in estimated power (scatter plot power minus overlaid line power) to reveal the true plot by smoothness and sample size with a joint 95\% confidence interval. The red line indicates the point where both plot designs are equally powerful The largest differences appear in negative correlation, where the overlaid line graph's power drops substantially.}
 \label{powerdiff}
\end{figure}

\begin{figure}[htbp]
  \centering
<<samplesize, dependson='setup', fig.width=6, fig.height=8, out.width='0.7\\textwidth', echo=FALSE, error=FALSE, message=FALSE, warning=FALSE>>=
ggplot(data=subset(bylineup, smoothed==0), aes(x=r, y=percent, colour=factor(n))) + geom_smooth(se=F) + geom_point() + facet_grid(design~.) + scale_colour_discrete(name = "Sample size") +  labs(y="Detection rate", x="Correlation") + theme(legend.position="bottom")
@
 \caption{Replotting figure~\ref{results} to highlight the effects of changes in sample size with unsmoothed data. The scatter plot appears to universally benefit from increased sample sizes but the overlaid lines have a mixed benefit, with improvements only in positive correlation.}
 \label{samplesize}
\end{figure}

@@

%The correlation:sign interaction shows that the scatter plot has close to symmetric performance for positive (correlation:sign + correlation) and negative (-correlation) correlation while the overlaid line graph is substantially weaker for negative values of correlation.
Smoothing out the excess 'noise' from a time series had a minimal effect on scatter plots, demonstrated by $\beta_{4} + \beta_{10}$, but it did have a significant positive impact on the power of overlaid line graphs as smoothing induces a large visual difference (See Figure~\ref{smoothing}. Increasing the sample size, $t$, improved the power of the scatter plot but it was not statistically significant for the overlaid line graph possibly due to the negative correlation results interfering with the modeling (See Figure~\ref{samplesize}).
